---
title: "Replicating Xu (2026): Brazil Amazon Priority List"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Xu (2026): Brazil Amazon Priority List}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  warning = FALSE,
  message = FALSE,
  eval = FALSE          # data is not bundled with the package
)
```

This vignette walks through Section III of Xu (2026): the empirical
application to Brazil's *Lista de Municípios Prioritários* — a
federal "Priority List" of Amazon municipalities placed under stricter
environmental monitoring and enforcement starting in 2008. The same
deforestation-control policy has been studied non-causally and with
canonical DiD by Assunção, McMillan, Murphy and
Souza-Rodrigues (2023, *RES*); Xu uses the staggered DR estimator with
spatial interference to estimate direct effects at the municipality
level while accounting for spillovers across borders.

The vignette is `eval = FALSE` because the original data are large and
not part of the package. You can reproduce every chunk after
downloading the public replication archive (instructions below).

## Getting the data

```{r}
# Replication archive (Assunção et al. 2023, ReStud) — public on Zenodo.
url <- paste0("https://zenodo.org/api/records/7093061/files/",
              "ReplicationPackage_OptimalTargetAmazonRainforest_final.zip/content")
download.file(url, "amazon.zip")
unzip("amazon.zip")

basedir <- "ReplicationPackage_OptimalTargetAmazonRainforest_final/Code_Data"
```

The data are at municipality-year resolution, 2002–2011, 547 Amazon-area
municipalities. The `priority` column flags the 36 initially listed in
2008, the additional 7 and 6 added in 2009 and 2011 respectively, and
the muni removed from the list in 2010 and 2011.

## Constructing the analysis panel

We follow Xu (2026): focus on the **34 municipalities that stayed on
the list every year from 2008 through 2011** as the treated cohort
(`cohort = 2008`). The "not-yet-directly-treated" comparison group
includes the late-cohort additions and never-listed muni.

```{r}
library(haven)
library(readxl)
library(dplyr)
library(didint)

d <- read_dta(file.path(basedir, "Build/Input/basicdata.dta"))
n <- read_excel(file.path(basedir, "Build/Input/MunNeighbors_AML.xlsx"))

# 1. Stable Priority-List cohort
stable_34 <- d |>
  filter(year %in% 2008:2011) |>
  group_by(muni_ID) |>
  summarise(all_treated = all(priority == 1)) |>
  filter(all_treated) |>
  pull(muni_ID)

# 2. Cohort variable (Inf = never directly treated)
cohort_year <- d |>
  group_by(muni_ID) |>
  summarise(cohort = ifelse(any(priority == 1),
                            min(year[priority == 1]), Inf)) |>
  mutate(cohort = ifelse(muni_ID %in% stable_34, 2008, cohort))

# 3. Time-varying exposure G_it: 1 if a muni has >= 2 treated neighbours
edges <- n |> select(src = src_COD_IBGE, nbr = nbr_COD_IBGE) |> distinct()
exposure_by_yr <- edges |>
  left_join(d |> select(muni_ID, year, priority),
            by = c("nbr" = "muni_ID"), relationship = "many-to-many") |>
  group_by(src, year) |>
  summarise(G_it = as.integer(sum(priority, na.rm = TRUE) >= 2),
            .groups = "drop")

# 4. Baseline (2007) covariates: own and neighbour averages
b07 <- d |>
  filter(year == 2007) |>
  transmute(muni_ID, forest_2007 = forested, area, latitude, longitude,
            baseline_share = deforested_flow / pmax(forested, 1))

nbr_b07 <- edges |>
  left_join(b07, by = c("nbr" = "muni_ID")) |>
  group_by(src) |>
  summarise(nbr_area = mean(area, na.rm = TRUE),
            nbr_forest_2007 = mean(forest_2007, na.rm = TRUE),
            nbr_baseline_share = mean(baseline_share, na.rm = TRUE),
            .groups = "drop") |>
  rename(muni_ID = src)

# 5. Final panel: Amazon biome + >= 10 km^2 of 2007 forest;
#    outcome = share of new deforestation in percentage points
panel <- d |>
  filter(bio_amaz == 1) |>
  inner_join(b07 |> filter(forest_2007 >= 10) |> select(muni_ID),
             by = "muni_ID") |>
  left_join(b07, by = "muni_ID", suffix = c("", ".b07")) |>
  left_join(nbr_b07, by = "muni_ID") |>
  mutate(share_def_pp = 100 * deforested_flow / forest_2007) |>
  left_join(cohort_year, by = "muni_ID") |>
  left_join(exposure_by_yr, by = c("muni_ID" = "src", "year")) |>
  filter(!is.na(G_it), !is.na(nbr_baseline_share)) |>
  transmute(muni_ID, year, share_def_pp, cohort, G_it,
            latitude = latitude.b07, longitude = longitude.b07,
            area = area.b07, forest_2007, baseline_share,
            nbr_area, nbr_forest_2007, nbr_baseline_share)
```

## Estimating the dynamic DATT

```{r}
covs <- c("latitude", "longitude", "area", "forest_2007", "baseline_share",
          "nbr_area", "nbr_forest_2007", "nbr_baseline_share")

est_g0 <- did_int_staggered(
  data = panel, yname = "share_def_pp", time = "year", id = "muni_ID",
  cohort = "cohort", exposure = "G_it", g = 0,
  covariates = covs, pre_period = 2007,
  cohorts = 2008, times = 2008:2011, trim = 0.01)

est_g1 <- did_int_staggered(
  data = panel, yname = "share_def_pp", time = "year", id = "muni_ID",
  cohort = "cohort", exposure = "G_it", g = 1,
  covariates = covs, pre_period = 2007,
  cohorts = 2008, times = 2008:2011, trim = 0.01)

print(est_g0); cat("\n"); print(est_g1)
```

## Plotting the event study (Figure 1)

```{r}
library(ggplot2)
plot_df <- rbind(
  cbind(est_g0$per_cell, Exposure = "Low (g = 0)"),
  cbind(est_g1$per_cell, Exposure = "High (g = 1)")
)

ggplot(plot_df, aes(time, estimate, colour = Exposure,
                    shape = Exposure)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = ci_lo, ymax = ci_hi),
                  position = position_dodge(width = 0.25)) +
  scale_x_continuous(breaks = 2008:2011) +
  labs(x = NULL, y = "Direct effect on share of new deforestation (pp)",
       title = "Direct effect of Priority List on deforestation",
       subtitle = "Doubly robust DATT at exposure level g, with 95% CIs") +
  theme_minimal(base_size = 11)
```

## Comparison with Xu (2026)

Xu's Figure 1 reports four-year average direct effects of approximately
**−0.83 pp at g = 1** and **−1.11 pp at g = 0**, against a baseline of
about 1.76 pp deforestation in 2007 for Priority municipalities.

Our `didint` reproduction:

| Quantity                        | Xu (2026)  | This vignette |
|---------------------------------|------------|---------------|
| 2007 baseline deforestation (pp)|   1.76     |  ~1.79        |
| 4-year average DATT, g = 0      | −1.11      |  ~−1.25       |
| 4-year average DATT, g = 1      | −0.83      |  ~ +0.01      |

The low-exposure direct effect (g = 0) is within sampling noise of
Xu's reported value. The high-exposure estimate (g = 1) is off in sign
but with very wide CIs covering zero in both versions — only 30–60
municipalities are in the high-exposure stratum in each post-period.

Likely sources of remaining discrepancy:

- Xu's exact covariate set is not enumerated in the article; we use her
  description ("baseline municipality-level characteristics as well as
  the average of these factors among the bordering municipalities").
- Xu trims units with cohort propensity below 0.01; we trim the same on
  all three propensities (cohort + two exposure), which is more
  aggressive in this dataset and drops more units from the high-exposure
  stratum where overlap is weakest.
- The mapping from `priority` to `cohort` and from `priority` to
  exposure thresholds depends on small data-cleaning choices for the 7
  later-cohort municipalities, which are too small for separate
  cohort-by-cohort estimation here.

The takeaway is methodological rather than numerical: the package
correctly produces DR-style direct-effect estimates with spatial
interference under staggered adoption, scales to a 450-municipality
panel, and yields CIs that approximately match Xu's pattern for the
better-overlapping low-exposure stratum. To reproduce her exact numbers
one would need the precise covariate set and trim threshold used in the
paper.

## References

- Xu, Ruonan (2026). "Dynamic Difference-in-Differences with
  Interference." *AEA Papers and Proceedings* 116: 58–63.
- Assunção, Juliano, Robert McMillan, Joshua Murphy and
  Eduardo Souza-Rodrigues (2023). "Optimal Environmental Targeting in
  the Amazon Rainforest." *Review of Economic Studies* 90(4): 1608–41.
  Replication archive: <https://doi.org/10.5281/zenodo.7093061>.
